---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LinearRegression

<!-- badges: start -->
[![R-CMD-check](https://github.com/yummymmz/Biostat625HW4/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yummymmz/Biostat625HW4/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/yummymmz/Biostat625HW4/graph/badge.svg)](https://app.codecov.io/gh/yummymmz/Biostat625HW4)
<!-- badges: end -->

`LinearRegression` is an R package that implements Ordinary Least Squares (OLS) regression using both R and C++ (via **Rcpp**) for improved computational efficiency.  
It provides lightweight and flexible functions for **model fitting**, **prediction**, 
and **model summarization**, mirroring the functionality of R’s built‑in `lm()` function.

This package is developed as part of **Biostat 625 Homework #4**, with the goal of 
demonstrating how to build an R package from scratch and compare the implementation 
against the original R base functions in terms of **correctness** (via `all.equal()`) 
and **efficiency** (via `microbenchmark()`).

Key features include:

### **Required components**
1. **A complete R package**, including well‑documented R source code (`R/`), function help pages 
   (`man/`), and working examples that run without errors or warnings.
2. **Public GitHub repository** hosting the package, with a clearly written `README.md` file 
   introducing its purpose, installation, and usage.
3. **Comprehensive vignette(s)** written in R Markdown that demonstrate the usage of the core 
   functions (e.g., model fitting, prediction, summary).
   The primary tutorial is provided in the vignette 
   [`vignettes/getting_started.Rmd`](vignettes/getting_started.Rmd).
4. **Comparison(s) against R's base `lm()` function** on simulated datasets are included in the 
   vignette [`vignettes/comparison.Rmd`](vignettes/comparison.Rmd), demonstrating both the  
   *numerical correctness* (checked with `all.equal()`) and the *computational efficiency* 
   (evaluated using `microbenchmark()`).

### **Optional bonus components**
1. **C++ integration via Rcpp**, providing a fast and efficient implementation of **QR decomposition** for    solving the Ordinary Least Squares (OLS) problem. The C++ function `lm_qr_extended_cpp()` is called        internally by `lm_fit()`.
2. **Unit testing** implemented with the **testthat** framework to verify correctness of all 
   major functions.
3. **Continuous integration (CI)** configured on GitHub using **GitHub Actions**, ensuring 
   that all tests are automatically executed for every commit and pull request.
4. **Code coverage monitoring** integrated via **Codecov**, verifying that all R and C++ code 
   paths are exercised during testing (100% coverage).


---

## Installation

You can install the development version of **LinearRegression** directly from GitHub:

```r
# install.packages("devtools")
devtools::install_github("yummymmz/Biostat625HW4",build_vignettes = TRUE)
```

After installation, load the package:

```r
library(LinearRegression)
```

Use ?lm_fit to get help page of the function. If it doesn't show, you can restart R and try it again.

---

## Example

Here is a short example that demonstrates how to fit and summarize a model
using `lm_fit()`, `summary.OLS()`, and `predict.OLS()`:

```{r example}
## Example 1: using built-in dataset (mtcars)
library(LinearRegression)
data(mtcars)

# Fit OLS model
fit_mtcars <- lm_fit(mpg ~ wt + hp, data = mtcars)

# Inspect results
fit_mtcars$coefficients

# Summarize
summary(fit_mtcars)

# Prediction
head(predict(fit_mtcars))

# Predict on new data
new_data <- data.frame(wt = c(2, 3), hp = c(120, 180))
predict(fit_mtcars, newdata = new_data)

## Example 2: using simulated data
set.seed(42)
df <- data.frame(
 x1 = rnorm(100),
 x2 = rnorm(100)
)
df$y <- 2 + 1.5*df$x1 - 3*df$x2 + rnorm(100, sd = 0.5)

# Fit OLS model
fit_sim <- lm_fit(y ~ x1 + x2, data = df)

# Inspect results
fit_sim$coefficients

# Summarize
summary(fit_sim)

# Prediction
predict(fit_sim)
```

---

## Learning more

A detailed tutorial and performance comparison with R’s `lm()` 
are included in the vignette `vignettes/getting_started.Rmd` and `vignettes/comparison.Rmd`.

```r
browseVignettes("LinearRegression")
```

---

## Citation

If you use **LinearRegression** in coursework or research, please cite this repository:
> Jiamin Zou (2025). *LinearRegression: Simple OLS Regression Implementation in R.*  
> GitHub repository: <https://github.com/yummymmz/Biostat625HW4>

---

## License

This package is released under the MIT License.
