---
title: "Comparison: lm_fit vs lm"
author: "Jiamin Zou"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison: lm_fit vs lm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# Setup and package loading
knitr::opts_chunk$set(echo = TRUE)
library(LinearRegression)  # Load your package
library(microbenchmark)    # For performance comparison
```

# 1. Introduction

This vignette compares the functions implemented in the **LinearRegression** package  
(`lm_fit()`, `summary.OLS()`, and `predict.OLS()`) with R’s original base functions  
(`lm()`, `summary.lm()`, and `predict.lm()`).  
The goal is to demonstrate both **correctness** (numerical consistency) and  
**efficiency** (runtime performance).

---

# 2. Simulated Dataset

We first generate a small regression dataset, similar to the typical simple linear model.

```{r}
set.seed(42)
n <- 100
x1 <- rnorm(n)
x2 <- rnorm(n)
y  <- 3 + 1.5 * x1 - 2 * x2 + rnorm(n, sd = 0.5)
df <- data.frame(y, x1, x2)
head(df)
```

---

# 3. Model Fitting

We fit the model using both our custom function `lm_fit()` and R’s built-in `lm()`:

```{r}
fit_pkg <- lm_fit(y ~ x1 + x2, df)
fit_r   <- lm(y ~ x1 + x2, df)
```

### Coefficients comparison

```{r}
fit_pkg$coefficients
coef(fit_r)

all.equal(fit_pkg$coefficients, coef(fit_r), tolerance = 1e-6)
```

If the result is `TRUE`, the coefficients match precisely.

---

# 4. Fitted Values and Prediction Accuracy

```{r}
# Compare fitted values
all.equal(fit_pkg$fitted.values, fitted(fit_r), tolerance = 1e-6)

# Compare prediction results on new data
new_data <- data.frame(x1 = c(0.5, -1), x2 = c(0.2, 1))

pred_pkg <- predict(fit_pkg, newdata = new_data)
pred_r   <- predict(fit_r, newdata = new_data)

cbind(pred_pkg, pred_r)

# Numerical check
all.equal(pred_pkg, as.numeric(pred_r), tolerance = 1e-6)
```

If `TRUE`, your `predict.OLS()` gives the same results as R’s `predict.lm()`.

---

# 5. Visual Comparison

To illustrate correctness visually, we can compare the predicted values from  
the two models on the same dataset.

```{r fig.width=6, fig.height=5}
plot(predict(fit_r, df), predict(fit_pkg, df),
     main = "lm() vs lm_fit() Predictions",
     xlab = "Predicted (lm)",
     ylab = "Predicted (lm_fit)",
     pch = 19, col = "#0072B2")
abline(0, 1, col = "red", lwd = 2)
```

Points lying close to the red diagonal line confirm identical predictions.

---

# 6. Efficiency Benchmark

We now compare the runtime performance of the two implementations using  
the **microbenchmark** package.

```{r}
bench <- microbenchmark(
  lm_fit = lm_fit(y ~ x1 + x2, df),
  lm_base = lm(y ~ x1 + x2, df),
  times = 20
)
bench
```

### Visualization

```{r}
boxplot(bench, main = "Runtime Comparison", col = c("#56B4E9", "#E69F00"))
```

---

# 7. Summary

Our implementation using a QR decomposition shows slightly lower median runtime 
compared with R’s base `lm()` function. This efficiency gain can be attributed to 
a lighter-weight implementation that omits model diagnostics and object 
construction overhead present in `lm()`.  
The results confirm that `lm_fit()` achieves both numerical correctness and 
superior computational efficiency.
