---
title: "Getting Started with LinearRegression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with LinearRegression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(LinearRegression)
set.seed(42)
```

# 1. Introduction

The `LinearRegression` package implements **Ordinary Least Squares (OLS)** regression using 
both R and C++ (via **RcppArmadillo**) for improved computational efficiency.  
This vignette demonstrates how to fit, summarize, and predict OLS models using the methods provided by this package.

You'll see:
- How to create or simulate data for regression analysis  
- How to fit a model with `lm_fit()`  
- How to examine results with `summary.OLS()`  
- How to generate predictions with `predict.OLS()`  

---

# 2. Example Data: Simulated Data Example

You can use either an **included sample dataset** or **simulated data**.


```{r simulate}
# Simulate regression data
set.seed(42)
n <- 100
x1 <- rnorm(n)
x2 <- rnorm(n)
# True model: y = 2 + 1.5*x1 - 3*x2 + error
y  <- 2 + 1.5*x1 - 3*x2 + rnorm(n, sd = 0.5)
df <- data.frame(y, x1, x2)
head(df)
```


---

# 3. Fitting a Model

We can fit a linear model using both our custom `lm_fit()` function 
and R’s built-in `lm()` for comparison.

```{r model-fit}
fit_pkg <- lm_fit(y ~ x1 + x2, data = df)
fit_r   <- lm(y ~ x1 + x2, data = df)
```

---

# 4. Summarizing Results

`summary(fit_pkg)` provides a detailed statistical summary of your fitted model, 
including coefficients, t-tests, R², adjusted R², residual SE, and F-statistic.

```{r model-summary}
summary(fit_pkg)
```

You can verify the correctness by comparing with base R’s output:

```{r model-summary-r}
summary(fit_r)
```

---

# 5. Making Predictions

The function `predict.OLS()` can generate both fitted and new predicted values.

```{r predictions}
# Fitted values (training)
head(predict(fit_pkg))

# New data prediction
new_df <- data.frame(x1 = c(0.5, -1), x2 = c(0.2, 1.0))
predict(fit_pkg, newdata = new_df)
```



# 6. Extension Example: Included Dataset Example (`mtcars`)


You can also try the built-in dataset `mtcars` in base R, for a familiar example.

```{r sample-data}
head(mtcars)
```

You can also analyze real datasets such as `mtcars`.

```{r mtcars-example}
model_mtcars <- lm_fit(mpg ~ wt + hp, data = mtcars)
summary(model_mtcars)
head(predict(model_mtcars))
```

---

# 7. Conclusion

This vignette demonstrates how to:
- simulate regression data or use built-in datasets,  
- fit an OLS model using `lm_fit()`,  
- summarize model results,  
- and make predictions.

Further comparison with R’s built-in `lm()` (accuracy and computational performance)
can be found in the vignette `comparison.Rmd`.
